### 2021-07-15 chapters/7_discussion.tex
finish manual reading notes

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering the mutants.
First we discuss the results of the first experiment. 
We will discuss the all the variant results together and then compare the results between the variants.
Second we will discuss the results of the second experiment and compare them to those of the first.
And last we will discuss the practicality of our research.

@@

66\% of the displayed results had a deviation of less than 1\%. 

@@

The results prove that the relations between the characteristics we identified can be used to cluster mutants efficiently. 
This can be explained by the correlation between the result of a mutant and the euclidean distance used to decide whether a mutants should be clustered together.

@@

    With a reduction up to 75\% of the total amount of mutants we can still maintain effectiveness.

@@

In both variants we observe a reduction in average accuracy when reducing the number of clusters.

@@

    The more you reduce the amount of mutants executed the more accuracy is lost.

@@

    We can reduce the amount of mutants executed in our data up to half of the original amount before reducing the accuracy even more.

@@

With the exception of Google Auto Service and Commons Text, with a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.

@@

As logic dictates we expect that a higher accuracy means a score closer to that of the original.

@@

As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance took a long time.
For this reason we repeated the experiment without the Levenshtein distance.
Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation score of.

@@

Clustering without Levenshtein distance is on average more accurate.

@@

    While clustering without the Levenshtein distance was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
The number of clusters compared to the that of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low number of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
\end{finding}

\section{Application}
The application of our research is limited to all software projects written in the language Java and the limitations of the selected mutation testing tool PIT\cite{pit}.
We can repeat our experiment for every java project that can be run by PIT.
This scope is significantly larger than most of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}).
The extraction of mutants is build within the PIT plugin system which is a plug and play system.
Applying our research can be done in three steps.
The first step is to run PIT with the plugin to gather the characteristics.
This plugin makes sure PIT exits without executing mutants after gathering the characteristics.
The second step is to run the script for either the machine learning model or the hierarchical clustering.
The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
It is possible to build all these steps into a DevOps pipeline to automate the process. 
The plugins and scripts needed for these steps are available on Github\cite{rbasarat-repo}.

\section{Threats to validity}
In this section, we discuss the threats to internal validity, external validity and reliability.
\subsection{Internal validity}
\subsubsection{Equivalent mutants}
Equivalent mutants are mutants that are syntactically different but functionally the same.
These mutants influence the mutation score while giving the same result as their equivalent counterpart.
In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
PIT contains some functionality for reducing equivalent mutants\cite{pit}.
We also reduced the chance of inaccuracy by repeating our experiment.

\subsubsection{Test set representativity}
The software projects we used in our research were systematically selected.
Chapter \ref{ch:project_selection} provides an indication but not a wide representation of industrial software projects.
The conclusions drawn regarding the increase performance by reducing the mutants are based on the comparison between our own baseline measurements.
Another threat which is the result of the representation of our test set is the representation of the mutants.
While our data set contains around 1 million mutants we do not know if we have seen all mutants that exist.
This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of PIT, ever written.

\subsubsection{Project configuration}
Some tests may not be recognized by PIT due to some external data required for the test.
Another reason may be that the test uses global variables that are not available to PIT.
We encountered some of these tests in our data set.
While these tests did pass when run with only the unit test framework they would not when run by PIT.
These projects still met the requirements stated in Chapter \ref{ch:project_selection}.
To solve this issue we removed these tests from the test suite.
By removing these specific tests we may have slightly, tough negligible, deviated from the individual real-world settings for respective projects.

>>>>

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiments on clustering mutants.
First, we discuss the results of the first experiment. 
We will discuss the all results together and then compare the results between the experiments including and excluding the Levenshtein distance characteristic.
Second we will discuss the results of the second experiment and compare them to those of the first.
And last we will discuss the practicality of our research.

@@

66\% of the displayed results have a deviation of less than 1\%. 

@@

The results prove that the relations between the mutants represented by the characteristics we identified can be used to cluster mutants efficiently. 
This can be explained by the correlation between the result of a mutant, the calculated euclidean distance and the distance threshold(linkage criteria) used to decide whether a mutants should be clustered together.

@@

    When executing 25\% of the total amount of mutants we can still maintain effectiveness.

@@

In both variations of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.

@@

    The more you reduce the number of mutants executed the more accuracy is lost.

@@

    We can reduce the number of mutants executed in our data set up to 50\% before losing more accuracy.

@@

With the exception of Google Auto Service and Commons Text, a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.

@@

According the rules of logic, a higher cluster accuracy means a score closer to that of the original.
In other words if all clusters had an accuracy of 100\% we will achieve the same mutation score as that of the full set.

@@

As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance characteristic took a long time.
For this reason we repeated the experiment without the Levenshtein distance characteristic.
Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation scores.

@@

Clustering without Levenshtein distance characteristic is on average more accurate.

@@

    While clustering without the Levenshtein distance characteristic was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance characteristic) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the \acrshort{fcm} model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
The number of clusters compared to the that of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the \acrshort{fcm} model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the \acrshort{fcm} model and hierarchical clustering we can observe that the lower accuracy of the \acrshort{fcm} model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low number of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
\end{finding}

\section{Application}
The application of our research is limited to all software projects written in the language Java and the requirements of PIT\cite{pit}.
We can repeat our experiment for every java project that can be run by PIT.
The scope of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}) is a subset of the scope in our research.
With our research it is possible to cluster all the mutants used in other related research.
The extraction of mutants is build within the PIT plugin system which is a plug and play system.
Applying our research can be done in three steps.
The first step is to run PIT with the plugin to gather the characteristics.
Gathering the characteristics does not take more than a few minutes depending on the number of mutants generated.
The plugin also makes sure PIT exits without executing mutants after gathering the characteristics.
The second step is to run the script for either the \acrshort{fcm} model or the hierarchical clustering.
The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
It is possible to build all these steps into a DevOps pipeline to automate the process. 
The plugins and scripts needed for these steps are available in the code repository on Github\cite{rbasarat-repo}.

\section{Threats to validity}
In this section, we discuss the threats to internal validity, external validity and reliability.
\subsection{Internal validity}
\subsubsection{Equivalent mutants}
Equivalent mutants are mutants that are syntactically different but functionally identical.
These mutants influence the mutation score while giving the same result as their equivalent counterpart.
In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
PIT contains some functionality for reducing equivalent mutants\cite{pit}.
We also reduced the chance of inaccuracy by repeating our experiment.

\subsubsection{Test set representativity}
The software projects we used in our research were systematically selected.
Chapter \ref{ch:project_selection} provides an indication but not a wide representation of industrial software projects.
The conclusions drawn regarding the increase in performance by reducing the mutants are based on the comparison between our own baseline measurements.
\newline
Another threat which is the result of the representation of our test set is the representation of the mutants.
While our data set contains around 1 million mutants we do not know if we have seen all mutants that exist.
This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of our research, ever written.

\subsubsection{Project configuration}
Some tests may not be recognized by PIT due to some external data required for the test.
Another reason may be that the tests uses global variables that are not available to PIT.
We encountered some of these tests in our data set.
While these tests did pass when run with only the unit test framework they would fail when run by PIT.
These projects still met the requirements stated in Chapter \ref{ch:project_selection}.
To solve this issue we removed these tests from the test suite.
By removing these specific tests we may have slightly, tough negligible, deviated from the individual real-world settings for respective projects.

### 2021-07-15 chapters/8_related_work.tex
finish manual reading notes

\subsection{Clustering by scope}
Yu et al., extended the research about overlapping mutants\cite{Yu2019PossibilityScope}.
They added more mutators to the scope.
The also extended the research to include statements and blocks.
Their results shows an increase in clustered mutants.
The problems mentioned in Chapter \ref{ch:overlapping_mutants} are not addressed in this research.

\subsection{Clustering Hamming distance}
Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
As a result they identified the Hamming distance for to cluster with. 
The Hamming distance is, like the Levenshtein distance, a similarity measure.
They use the k-means to cluster the mutants with the Hamming distance.
They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
But they do remark that their research still has problems with rationally determining the domains of their clusters.
Our research does not have this problem.
In our reserach the centre of of the domains are decided by the characteristics of the mutants.

\subsection{Spectral clustering}
Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
With these definitions they reduce the mutants based on their proposed reduction method.
This method is based on spectral clustering (SCMT), including the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
The reduced were then clustered with a classical clustering algorithm.
Their results are promising and show good cluster accuracy.
They write that there is still work left to do in optimizing the matrices and clustering algorithms.
Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.

\subsection{Clustering similarity}
Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
They used the Hamming distance a similarity measure.
They calculated the distance and used this as data te feed into the clustering algorithms.
The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
They did take into account efficiency.
Hussain et al, had the same goals as our research. 
The difference is that our research used different and more elaborated methods to cluster the mutants.

\subsection{Generalizing mutants}
Wilinski et al., tried to generalize the mutants by defining metrics.
Each metric is calculated for a specific mutator.
The three metrics are usefulness, frequency and dependency.
Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable mutant cost reduction without a loss of the mutation score accuracy.
Wilinski et al., their research is very narrowly scoped to the specific mutators they decided to do research on.
Our research does not contain this limitation.
The mutator is used as a characteristic in our research instead of limiting our research it helps define the the mutant.

>>>>

\subsection{Clustering by scope}
Yu et al., extended the research about overlapping mutants(See chapter \ref{ch:overlapping_mutants})\cite{Yu2019PossibilityScope}.
They extended the scope by adding more mutators.
The also extended the research to include statements and blocks.
Their results shows an increase in clustered mutants.
The research of Yu et al., suffers from the same limitations and threats  mentioned in Chapter \ref{ch:overlapping_mutants}.

\subsection{Clustering Hamming distance}
Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
As a result they identified the Hamming distance to cluster with. 
The Hamming distance is, like the Levenshtein distance, a similarity measure.
They use the k-means algorithm to cluster mutants represented by Hamming distance.
They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
\newline
Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
They do acknowledge that their research still has problems with rationally determining the domains of their clusters.
Our research does not have this problem, as the centre of the domains are decided by the characteristics of the mutants.

\subsection{Spectral clustering}
Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
With these definitions they reduced the mutants based on their proposed reduction method.
This method is based on spectral clustering (SCMT), the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
The reduced mutants were then clustered with a classical clustering algorithm.
Their results are promising and show high cluster accuracy.
They write that there is still work left to do in optimizing the matrices and clustering algorithms.
Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.
Another difference is that our research does not require to know the result of a mutant(killed or survived).

\subsection{Clustering similarity}
Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
They used the Hamming distance as similarity measure.
They calculated the distance and used this as data to feed into the clustering algorithms.
The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
They did take into account efficiency.
Hussain et al, had the same goals as our research. 
The difference is that our research used different and more elaborated methods to cluster the mutants.
Hussain et al., calculated the mutation score by counting one cluster as one mutant.
Our research calculated a weighted mutation score to reflect the clusters more accurately.

\subsection{Generalizing mutants}
Wilinski et al., tried to generalize the mutants by defining metrics.
Each metric is calculated for a specific mutator.
The three metrics are usefulness, frequency and dependency.
Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable cost reduction without a loss of the mutation score accuracy.
Wilinski et al., their research is narrowly scoped to the specific mutators they decided to do research on.
Our research does not contain this limitation.
The mutator is used as a characteristic in our research instead of limiting our research it helps define the the mutant.
